Unsupervised embedding of trajectories captures the latent structure of mobility

### 2020-10-19 paper/current/main.tex
first main para

We show that the embedding distance better predicts actual mobility flows than the geographic distance.
In the case flight itineraries, the embedding distance explains more than twice the expected flux between airports ($R^{2} = 0.51$, Fig.~\ref{fig:gravity_performance}a) than does geographic distance ($R^{2} = 0.22$). 
We fit the gravity model given by Eq.~\ref{eq:gravity_basic} (see Methods) and compute the expected flux $\hat{T}_{ij}$ using the geographic distance and the embedding distance as parameters.
The embedding distance produces better predictions of actual flux between airports than does the geographic distance (Fig.~\ref{fig:gravity_performance}b).
The neural embedding also performs well in a smaller geographic context:
In the case of South Korean hotel accommodations, such that the embedding distance better explains the expected flux ($R^{2} = 0.57$, Fig.~\ref{fig:gravity_performance}c) than does geographic distance ($R^{2} = 0.25$), and predictions made using embedding distance outperform those made with geographic distance (Fig.~\ref{fig:gravity_performance}d). 
This performance is consistent in the case of scientific mobility:
the embedding distance explains more than twice the expected flux ($R^{2} = 0.48$, Fig.~\ref{fig:gravity_performance}e) than does the geographic distance ($R^{2} = 0.22$), and predictions made using the embedding distance outperform those using the geographic distance (Fig.~\ref{fig:gravity_performance}f).
These patterns hold for the subsets of only domestic (within-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}c and Fig.~\ref{fig:supp:predict_geo_bymodel}) and only international mobility flows (across-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}d and  Fig.~\ref{fig:supp:predict_emb_bymodel}).
For scientific mobility, the embedding distance defined from the cosine distance between vectors also outperforms that defined from their dot product ($R^{2} = 0.34$, Fig.~\ref{fig:supp:gravity_dot}).
The embedding distance also out-performs alternative ``diffusion-based'' network distance measures such as the personalized-page rank scores (see \SI) calculated from the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
Our results demonstrate that the embedding distance better captures patterns of actual mobility than does the geographic distance across domains.

>>>>

We show that the embedding distance better predicts actual mobility flows than the geographic distance.
In the case of flight itineraries, the embedding distance explains more than twice the expected flux between airports ($R^{2} = 0.51$, Fig.~\ref{fig:gravity_performance}a) than does geographic distance ($R^{2} = 0.22$).  Also, the embedding distance produces better predictions of actual flux between airports than does the geographic distance (Fig.~\ref{fig:gravity_performance}b).
In the case of South Korean hotel accommodations, such that the embedding distance better explains the expected flux ($R^{2} = 0.57$, Fig.~\ref{fig:gravity_performance}c) than does geographic distance ($R^{2} = 0.25$), and predictions made using embedding distance outperform those made with geographic distance (Fig.~\ref{fig:gravity_performance}d). 
This performance is consistent in the case of scientific mobility:
the embedding distance explains more than twice the expected flux ($R^{2} = 0.48$, Fig.~\ref{fig:gravity_performance}e) than does the geographic distance ($R^{2} = 0.22$), and predictions made using the embedding distance outperform those using the geographic distance (Fig.~\ref{fig:gravity_performance}f).
These patterns hold for the subsets of only domestic (within-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}c and Fig.~\ref{fig:supp:predict_geo_bymodel}) and only international mobility flows (across-country organization pairs, Fig.~\ref{fig:supp:gravity_correlation}d and  Fig.~\ref{fig:supp:predict_emb_bymodel}).
The embedding distance also out-performs alternative diffusion-based network distance measures including the personalized-page rank scores  calculated from the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
In sum, our results demonstrate that the embedding distance better captures patterns of actual mobility than does the geographic distance consistently.
